Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-03-02 paper/main.tex
paper: some Impl & Experiments text + TeX for tables

@@

@@

@@

We have implemented our reduction technique in a tool called
\texttt{telatko}. We show that the tool can reduce acceptance marks in
automata produced by Delag~\cite{muller.17.gandalf},
ltl3tela~\cite{major.19.atva}, Rabinizer~4~\cite{kretinsky.18.cav},
and Spot~\cite{duret.16.atva2}.  While the reduction is relatively
modest for TELA produced by ltl3tela and Spot, it is substantial for
automata produced by Rabinizer~4. \todo{A jak vyjde Delag? Fixnout i
  Abstract.}

@@

  \caption{An example illustrating the results of the three basic
    reduction levels: an input automaton and the automata
    obtained by reduction of level 1, level 2, and level 3.
    \todo{Pozor, prvni klauzule v levelu 2 je trash!}}
  \label{fig:example}
\end{figure}

@@

>>>>

\usepackage{multirow}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{makecell}
\usepackage[table]{xcolor}
%\usepackage{arydshln}
%\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}

@@

\usepackage{pgfplots}
\pgfplotsset{compat=newest}

@@

\newcommand{\telatko}{\texttt{telatko}\xspace}

@@

We have implemented our reduction technique in a tool called
\telatko. We show that the tool can reduce acceptance marks in
automata produced by Delag~\cite{muller.17.gandalf},
ltl3tela~\cite{major.19.atva}, Rabinizer~4~\cite{kretinsky.18.cav},
and Spot~\cite{duret.16.atva2}.  While the reduction is relatively
modest for TELA produced by ltl3tela and Spot, it is substantial for
automata produced by Rabinizer~4. \todo{A jak vyjde Delag? Fixnout i
  Abstract.}

@@

  \caption{An example illustrating the results of the three basic
    reduction levels: an input automaton and the automata
    obtained by reduction of level 1, level 2, and level 3.
    \todo{Pozor, prvni klauzule v levelu 2 je trash!}}
  \label{fig:example}
\end{figure}
%}}}
%{{{ algorithm and implementation

\section{Reduction algorithm and its implementation}\label{sec:algimpl}

The presented reduction algorithms have been implemented in a tool called
\telatko. It uses the Spot library~\cite{duret.16.atva2} for automata parsing
and manipulation, and the Z3 Theorem Prover~\cite{demoura.08.tacas} to solve
the satisfiability of QBF formulas. Moreover, in the experimental evaluation
we use several tools from the Spot library, i.e. \texttt{genltl},
\texttt{ltlfilt}, \texttt{ltl2tgba}, \texttt{autcross} and \texttt{autfilt},
to generate, simplify and translate LTL formulas and to compare and simplify
automata. \telatko is implemented in Python.

Jak se nastavuji inicialni parametry a jak postupne klesaji.
Jak prechazim od nizsich levelu k vyssim.
Vyuzivame incremental QBF solving?

@@

We have used the tool \texttt{genltl} from Spot to generate LTL formulas
from literature and from scalable patterns. Using the values from 1 to 5 for
the parametrized formulas, we have generated 407 unique LTL formulas. These
have been translated to deterministic or nondeterministic TELA using four
state-of-the-art translators \texttt{ltl2tgba} (from the Spot library),
\texttt{ltl3tela}, \texttt{ltl2dela} (known as Delag), and \texttt{ltl2dgra}
(known as Rabinizer 4). The generated automata sets contain less than 407
automata as not all of the formulas are successfully translated due to timeouts.
Table \ref{tab:ltltranslators} shows the exact versions, the homepage of the
used tools and the number of input automata produced by the given tool from
the LTL formulas.

Spot version 2.10.4 and Z3 version 4.8.13 have been used in the experimental
evaluation.

\begin{table}
\caption{LTL to automata translators used for the experimental evaluation
and the number of successfully translated formulas.}
\label{tab:ltltranslators}
\centering
\begin{tabular}{lr@{~~~}lr}
\toprule
tool        & version&homepage & \# formulas \\
\midrule
\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 404\\
\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} \todo{RELEASE} & 376\\
Delag & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 397\\
Rabinizer & \todo ? & & 386\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\caption{Test: heatmap}
\centering
\begin{tabular}{c|ccccc|}
\multicolumn{1}{c}{} & 3 & 4 & 5 & 6 & \multicolumn{1}{c}{7} \\
%\specialrule{2pt}{0pt}{0pt}
\cline{2-6}
3 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \cellcolor{blue!25}\makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
4 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
5 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
6 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
7 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
\cline{2-6}
\end{tabular}
\end{table}

\begin{table}
\caption{Test: Data}
\centering
\begin{tabular}{cccccc}
\toprule
& & marks & \% red. & time & timeouts \\
\midrule
\multirow{4}{*}{\rotatebox[origin=c]{90}{\makecell{Delag \\ (420 marks)}}} & level 1 & 240 & 12.3 & 133.7 & 42 \\
& level 2 & 240 & 12.3 & 133.7 & 42 \\
& level 3 & 240 & 12.3 & 133.7 & 42 \\
& multilevel & 240 & 12.3 & 133.7 & 42 \\
\midrule
\multirow{4}{*}{\rotatebox[origin=c]{90}{\makecell{Rabinizer 4 \\ (420 marks)}}} & level 1 & 240 & 12.3 & 133.7 & 42 \\
& level 2 & 240 & 12.3 & 133.7 & 42 \\
& level 3 & 240 & 12.3 & 133.7 & 42 \\
& multilevel & 240 & 12.3 & 133.7 & 42 \\
\bottomrule
\end{tabular}
\end{table}

\begin{tikzpicture}
\begin{axis}[
	ymode=log,
	height=5cm,
	width=10cm,
	grid=major,
]
\addplot coordinates {
	(1, 0.3)
	(2, 0.8)
	(3, 1.5)
	(4, 2.7)
	(5, 3.9)
	(6, 4.1)
};
\addlegendentry{Delag}
\end{axis}
\end{tikzpicture}
